---
title: "Simulating Continuous-Time Markov Chains with simmer (2)"
author: "Iñaki Ucar"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
library(simmer)
create_trajectory <- trajectory
```

<script type="text/x-mathjax-config">
  MathJax.Ajax.config.path["Contrib"] = "https://cdn.mathjax.org/mathjax/contrib";
  MathJax.Hub.Config({
    TeX: {extensions: ["[Contrib]/xyjax/xypic.js","AMSmath.js","AMSsymbols.js"]},
    tex2jax: {inlineMath: [["$","$"],["\\(","\\)"]]}
  });
</script>

In part one, [we simulated a simple CTMC](http://fishyoperations.com/2016/04/15/simulating-ctmc.html). Now, let us complicate things a bit. Remember the example problem there:

> A gas station has a single pump and no space for vehicles to wait (if a vehicle arrives and the pump is not available, it leaves). Vehicles arrive to the gas station following a Poisson process with a rate of $\lambda=3/20$ vehicles per minute, of which 75% are cars and 25% are motorcycles. The refuelling time can be modelled with an exponential random variable with mean 8 minutes for cars and 3 minutes for motorcycles, that is, the services rates are $\mu_\mathrm{c}=1/8$ cars and $\mu_\mathrm{m}=1/3$ motorcycles per minute respectively (note that, in this context, $\mu$ is a rate, not a mean).

Consider the previous example, but, this time, __there is space for one motorcycle to wait__ while the pump is being used by another vehicle. In other words, cars see a queue size of 0 and motorcycles see a queue size of 1.

The new Markov chain is the following:

$$\vcenter{\xymatrix{
*=<15mm,8mm>[o][F]{car+} \ar@(r,u)[drr]^{\mu_\mathrm{c}} \\
*=<15mm,8mm>[o][F]{car} \ar@/_/[r]_{\mu_\mathrm{c}} \ar@/^/[u]^{(1-p)\lambda} &
*=<15mm,8mm>[o][F]{empty} \ar@/_/[l]_{p\lambda} \ar@/^/[r]^{(1-p)\lambda} &
*=<15mm,8mm>[o][F]{m/cycle} \ar@/^/[l]^{\mu_\mathrm{m}} \ar@/_/[d]_{(1-p)\lambda} \\
& & *=<15mm,8mm>[o][F]{m/c+} \ar@/_/[u]_{\mu_\mathrm{m}}
}}$$

$$Q = \begin{pmatrix}
-\mu_\mathrm{c} & 0 & 0 & \mu_\mathrm{c} & 0 \\
(1-p)\lambda & -(1-p)\lambda-\mu_\mathrm{c} & \mu_\mathrm{c} & 0 & 0 \\
0 & p\lambda & -\lambda & (1-p)\lambda & 0 \\
0 & 0 & \mu_\mathrm{m} & -(1-p)\lambda-\mu_\mathrm{m} & (1-p)\lambda \\
0 & 0 & 0 & \mu_\mathrm{m} & -\mu_\mathrm{m} \\
\end{pmatrix}$$

where the states _car+_ and _m/c+_ represent _car + waiting motorcycle_ and _motorcycle + waiting motorcycle_ respectively.

With $p$ the steady state distribution, the average number of vehicles in the system is given by

$$N = 2(p_1 + p_5) + p_2 + p_4$$

```{r}
# Arrival rate
lambda <- 3/20
# Service rate (cars, motorcycles)
mu <- c(1/8, 1/3)
# Probability of car
p <- 0.75

# Theoretical resolution
A <- matrix(c(1,                   0,       0,               mu[1],            0,
              1, -(1-p)*lambda-mu[1],   mu[1],                   0,            0,
              1,            p*lambda, -lambda,        (1-p)*lambda,            0,
              1,                   0,   mu[2], -(1-p)*lambda-mu[2], (1-p)*lambda,
              1,                   0,       0,               mu[2],       -mu[2]), 
            byrow=T, ncol=5)
B <- c(1, 0, 0, 0, 0)
P <- solve(t(A), B)
N_average_theor <- sum(P * c(2, 1, 0, 1, 2)) ; N_average_theor
```

As in the previous post, we can simulate this chain by breaking down the problem into two trajectories (one for each type of vehicle and service rate) and two generators. But in order to disallow cars to stay in the pump's queue, we need to introduce a little trick in the cars' `seize`: the argument `amount` is a function that returns 1 if the pump is vacant and 2 otherwise. This implies that the car gets rejected, because there is only one position in queue and that `seize` is requesting two positions. Note also that the environment `env` must be defined __before__ running the simulation, as it is needed inside the trajectory.

```{r}
library(simmer)
set.seed(1234)

option.1 <- function(t) {
  car <- create_trajectory() %>%
    seize("pump", amount=function() {
      if (env %>% get_server_count("pump")) 2  # rejection
      else 1                                   # serve
    }) %>%
    timeout(function() rexp(1, mu[1])) %>%
    release("pump", amount=1)
  
  mcycle <- create_trajectory() %>%
    seize("pump", amount=1) %>%
    timeout(function() rexp(1, mu[2])) %>%
    release("pump", amount=1)
  
  env <- simmer() %>%
    add_resource("pump", capacity=1, queue_size=1) %>%
    add_generator("car", car, function() rexp(1, p*lambda)) %>%
    add_generator("mcycle", mcycle, function() rexp(1, (1-p)*lambda))
  env %>% run(until=t)
}
```

The same idea using a branch, with a single generator and a single trajectory.

```{r}
option.2 <- function(t) {
  vehicle <- create_trajectory() %>%
    branch(function() sample(c(1, 2), 1, prob=c(p, 1-p)), c(F, F), 
           create_trajectory("car") %>%
             seize("pump", amount=function() {
               if (env %>% get_server_count("pump")) 2  # rejection
               else 1                                   # serve
             }) %>%
             timeout(function() rexp(1, mu[1])) %>%
             release("pump", amount=1),                 # always 1
           create_trajectory("mcycle") %>%
             seize("pump", amount=1) %>%
             timeout(function() rexp(1, mu[2])) %>%
             release("pump", amount=1))
  
  env <- simmer() %>%
    add_resource("pump", capacity=1, queue_size=1) %>%
    add_generator("vehicle", vehicle, function() rexp(1, lambda))
  env %>% run(until=t)
}
```

We may also avoid messing up things with branches and subtrajectories. We can decide the type of vehicle and set it as an attribute of the arrival with `set_attribute`. Then, every activity's function is able to retrieve those attributes as a named list. Although the `branch` option is a little bit faster, this one is nicer, because there are no subtrajectories involved.

```{r}
option.3 <- function(t) {
  vehicle <- create_trajectory("car") %>%
    set_attribute("vehicle", function() sample(c(1, 2), 1, prob=c(p, 1-p))) %>%
    seize("pump", amount=function(attrs) {
      if (attrs["vehicle"] == 1 &&
          env %>% get_server_count("pump")) 2    # car rejection
      else 1                                     # serve
    }) %>%
    timeout(function(attrs) rexp(1, mu[attrs["vehicle"]])) %>%
    release("pump", amount=1)                    # always 1
  
  env <- simmer() %>%
    add_resource("pump", capacity=1, queue_size=1) %>%
    add_generator("vehicle", vehicle, function() rexp(1, lambda))
  env %>% run(until=t)
}
```

But if performance is a requirement, we can play cleverly with the resource's capacity and queue size, and with the amounts requested in each `seize`, in order to model the problem without checking the status of the resource. Think about this:

* A resource with `capacity=3` and `queue_size=2`.
* A car always tries to seize `amount=3`.
* A motorcycle always tries to seize `amount=2`.

In these conditions, we have the following possibilities:

* Pump empty.
* One car (3 units) in the server [and optionally one motorcycle (2 units) in the queue].
* One motorcycle (2 units) in the server [and optionally one motorcycle (2 units) in the queue].

Just as expected! So, let's try:

```{r}
option.4 <- function(t) {
  vehicle <- create_trajectory() %>%
    branch(function() sample(c(1, 2), 1, prob=c(p, 1-p)), c(F, F), 
           create_trajectory("car") %>%
             seize("pump", amount=3) %>%
             timeout(function() rexp(1, mu[1])) %>%
             release("pump", amount=3),
           create_trajectory("mcycle") %>%
             seize("pump", amount=2) %>%
             timeout(function() rexp(1, mu[2])) %>%
             release("pump", amount=2))
  
  simmer() %>%
    add_resource("pump", capacity=3, queue_size=2) %>%
    add_generator("vehicle", vehicle, function() rexp(1, lambda)) %>%
    run(until=t)
}
```

We are still wasting time in the `branch` decision. We can mix this solution above with the `option.1` to gain extra performance:

```{r}
option.5 <- function(t) {
  car <- create_trajectory() %>%
    seize("pump", amount=3) %>%
    timeout(function() rexp(1, mu[1])) %>%
    release("pump", amount=3)
  
  mcycle <- create_trajectory() %>%
    seize("pump", amount=2) %>%
    timeout(function() rexp(1, mu[2])) %>%
    release("pump", amount=2)
  
  simmer() %>%
    add_resource("pump", capacity=3, queue_size=2) %>%
    add_generator("car", car, function() rexp(1, p*lambda)) %>%
    add_generator("mcycle", mcycle, function() rexp(1, (1-p)*lambda)) %>%
    run(until=t)
}
```

Options 1, 2 and 3 are slower, but they give us the correct numbers, because the parameters (capacity, queue size, amounts) in the model remain unchanged compared to the problem. For instance,

```{r}
gas.station <- option.1(5000)

library(ggplot2)

# Evolution + theoretical value
graph <- plot_resource_usage(gas.station, "pump", items="system")
graph + geom_hline(yintercept=N_average_theor)
```

However, it is not the case in options 4 and 5. The parameters of these models have been _adulterated_ to fit our performance purposes. Therefore, we need to extract the RAW data, rescale the numbers and plot them. And, of course, we get the same figure:

```{r, message=FALSE}
gas.station <- option.5(5000)

limits <- data.frame(item = c("queue", "server", "system"), value = c(1, 1, 2))

library(dplyr); library(tidyr)

graph <- gas.station %>% get_mon_resources() %>% 
  gather(item, value, server, queue, system) %>%
  mutate(value = round(value * 2/5),                  # rescaling here <------
         item = factor(item)) %>%
  filter(item %in% "system") %>%
  group_by(resource, replication, item) %>%
  mutate(mean = c(0, cumsum(head(value, -1) * diff(time))) / time) %>% 
  ungroup() %>%
  ggplot() + aes(x=time, color=item) +
  geom_line(aes(y=mean, group=interaction(replication, item))) +
  ggtitle("Resource usage: pump") +
  ylab("in use") + xlab("time") + expand_limits(y=0) +
  geom_hline(aes(yintercept=value, color=item), limits, lty=2)
graph + geom_hline(yintercept=N_average_theor)
```

Finally, these are some performance results:

```{r}
library(microbenchmark)

t <- 1000/lambda
tm <- microbenchmark(option.1(t), 
                     option.2(t), 
                     option.3(t),
                     option.4(t),
                     option.5(t))
graph <- autoplot(tm)
graph + scale_y_log10(breaks=function(limits) pretty(limits, 5)) +
  ylab("Time [milliseconds]")
```
