---
title: "Simulating queueing systems with simmer"
author: "Iñaki Ucar"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
library(simmer)
create_trajectory <- trajectory
```

We are very pleased to announce that a new release of **simmer**, the Discrete-Event Simulator for R, is on CRAN. There are quite a few changes and fixes, with the **support of preemption** as a star new feature. Check out the complete set of release notes [here](https://github.com/Enchufa2/simmer/releases/tag/v3.2.0).

Let's _simmer_ for a bit and see how this package can be used to simulate queueing systems in a very straightforward way.

## The M/M/1 system

In [Kendall's notation](https://en.wikipedia.org/wiki/Kendall's_notation), an M/M/1 system has exponential arrivals (__M__/M/1), a single server (M/M/__1__) with exponential service time (M/__M__/1) and an inifinite queue (implicit M/M/1/__$\infty$__). For instance, people arriving at an ATM at rate $\lambda$, waiting their turn in the street and withdrawing money at rate $\mu$.

Let us remember the basic parameters of this system:

$$\begin{aligned}
\rho &= \frac{\lambda}{\mu} &&\equiv \mbox{Server utilization} \\
N &= \frac{\rho}{1-\rho} &&\equiv \mbox{Average number of customers in the system (queue + server)} \\
T &= \frac{N}{\lambda} &&\equiv \mbox{Average time in the system (queue + server) [Little's law]} \\
\end{aligned}$$

whenever $\rho < 1$. If that is not true, it means that the system is unstable: there are more arrivals than the server is capable of handling, and the queue will grow indefinitely.

The simulation of an M/M/1 system is quite simple using `simmer`. The trajectory-based design, combined with `magrittr`'s pipe, is very _verbal_ and self-explanatory.

```{r}
library(simmer)
set.seed(1234)

lambda <- 2
mu <- 4
rho <- lambda/mu # = 2/4

mm1.trajectory <- create_trajectory() %>%
  seize("resource", amount=1) %>%
  timeout(function() rexp(1, mu)) %>%
  release("resource", amount=1)

mm1.env <- simmer() %>%
  add_resource("resource", capacity=1, queue_size=Inf) %>%
  add_generator("arrival", mm1.trajectory, function() rexp(1, lambda)) %>% 
  run(until=2000)
```

Our package provides convenience plotting functions to quickly visualise the usage of a resource over time, for instance. Down below, we can see how the simulation converges to the theoretical average number of customers in the system.

```{r}
library(ggplot2)

# Evolution of the average number of customers in the system
graph <- plot_resource_usage(mm1.env, "resource", items="system")

# Theoretical value
mm1.N <- rho/(1-rho)
graph + geom_hline(yintercept=mm1.N)
```

It is possible also to visualise, for instance, the instantaneous usage of individual elements by playing with the parameters `items` and `steps`.

```{r, warning=FALSE}
plot_resource_usage(mm1.env, "resource", items=c("queue", "server"), steps=TRUE) +
  xlim(0, 20) + ylim(0, 4)
```

We may obtain the time spent by each customer in the system and we compare the average with the theoretical expression.

```{r}
mm1.arrivals <- get_mon_arrivals(mm1.env)
mm1.t_system <- mm1.arrivals$end_time - mm1.arrivals$start_time

mm1.T <- mm1.N / lambda
mm1.T ; mean(mm1.t_system)
```

It seems that it matches the theoretical value pretty well. But of course we are picky, so let's take a closer look, just to be sure (and to learn more about `simmer`, why not). Replication can be done with standard R tools:

```{r}
library(parallel)

envs <- mclapply(1:1000, function(i) {
  simmer() %>%
    add_resource("resource", capacity=1, queue_size=Inf) %>%
    add_generator("arrival", mm1.trajectory, function() rexp(1, lambda)) %>%
    run(1000/lambda) %>%
    wrap()
})
```

_Et voilà!_ Parallelizing has the shortcoming that we lose the underlying C++ objects when each thread finishes, but the `wrap` function does all the magic for us retrieving the monitored data. Let's perform a simple test:

```{r, message=FALSE}
library(dplyr)

t_system <- get_mon_arrivals(envs) %>%
  mutate(t_system = end_time - start_time) %>%
  group_by(replication) %>%
  summarise(mean = mean(t_system))

t.test(t_system$mean)
```

Good news: the simulator works. Finally, an M/M/1 satisfies that the distribution of the time spent in the system is, in turn, an exponential random variable with average $T$.

```{r}
qqplot(mm1.t_system, rexp(length(mm1.t_system), 1/mm1.T))
abline(0, 1, lty=2, col="red")
```

## M/M/c/k systems

An M/M/c/k system keeps exponential arrivals and service times, but has more than one server in general and a finite queue, which often is more realistic. For instance, a router may have several processor to handle packets, and the in/out queues are necessarily finite.

This is the simulation of an M/M/2/3 system (2 server, 1 position in queue). Note that the trajectory is identical to the M/M/1 case.

```{r}
lambda <- 2
mu <- 4

mm23.trajectory <- create_trajectory() %>%
  seize("server", amount=1) %>%
  timeout(function() rexp(1, mu)) %>%
  release("server", amount=1)

mm23.env <- simmer() %>%
  add_resource("server", capacity=2, queue_size=1) %>%
  add_generator("arrival", mm23.trajectory, function() rexp(1, lambda)) %>%
  run(until=2000)
```

In this case, there are rejections when the queue is full.

```{r}
mm23.arrivals <- get_mon_arrivals(mm23.env)
mm23.arrivals %>%
  summarise(rejection_rate = sum(!finished)/length(finished))
```

Despite this, the time spent in the system still follows an exponential random variable, as in the M/M/1 case, but the average has dropped.

```{r}
mm23.t_system <- mm23.arrivals$end_time - mm23.arrivals$start_time
# Comparison with M/M/1 times
qqplot(mm1.t_system, mm23.t_system)
abline(0, 1, lty=2, col="red")
```
