---
title: "Extensions for 'simmer'"
author: "Iñaki Ucar"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A new version of the [Discrete-Event Simulator for R](http://r-simmer.org/) was released a few days ago on CRAN. The most interesting new feature is the implementation of the subsetting operators `[` and `[[` for trajectory objects. Basically, think about trajectories as lists of activities and these operators will do (almost) everything you expect.

```{r}
library(simmer)

t0 <- trajectory() %>%
  seize("resource", 1) %>%
  timeout(function() rexp(1, 2)) %>%
  release("resource", 2)

t0

t0[c(3, 1)]
```

After the last maintenance update (v3.5.1), which fixed several bugs and included [a new interesting vignette with SimPy examples](http://r-simmer.org/articles/E-simpy.html) translated to 'simmer', this v3.6.0 comes hand in hand with the first 'simmer' extension released on CRAN: [**simmer.plot**](http://r-simmer.org/extensions/plot/). 

The primary purpose of 'simmer.plot' is to detach plotting capabilities from the core package, to systematise and enhance them. If you were using any of the old `plot_*()` functions, you will get a deprecation warning pointing to the S3 method `simmer.plot::plot.simmer`. [This vignette](http://r-simmer.org/extensions/plot/articles/plot.simmer.html) will help you make the transition.

'simmer.plot' also implements a new `plot` S3 method for trajectories. It produces a diagram of a given trajectory object, which is very helpful for debugging and checking that everything conforms your simulation model. Let us consider, for instance, the following pretty complex trajectory:

```{r}
t0 <- trajectory() %>%
  seize("res0", 1) %>%
  branch(function() 1, c(TRUE, FALSE),
         trajectory() %>%
           clone(2,
                 trajectory() %>%
                   seize("res1", 1) %>%
                   timeout(1) %>%
                   release("res1", 1),
                 trajectory() %>%
                   trap("signal",
                        handler=trajectory() %>%
                          timeout(1)) %>%
                   timeout(1)),
         trajectory() %>%
           set_attribute("dummy", 1) %>%
           seize("res2", function() 1) %>%
           timeout(function() rnorm(1, 20)) %>%
           release("res2", function() 1) %>%
           release("res0", 1) %>%
           rollback(11)) %>%
  synchronize() %>%
  rollback(2) %>%
  release("res0", 1)
```

We must ensure that:

- Resources are seized and released as we expect.
- Branches end (or continue) where we expect.
- Rollbacks point back to the activity we expect.
- ...

Things are indeed much easier if you can just inspect it visually:

```{r, message=FALSE}
library(simmer.plot)

plot(t0)
```

Note that different resources are mapped to a qualitative color scale, so that you can quickly glance whether you placed the appropriate seizes/releases for each resource.

Other interesting 'simmer' extensions are already on our roadmap. Particularly, Bart has been _simmering_ a new package (still under development) called [**simmer.optim**](https://github.com/r-simmer/simmer.optim), which brings parameter optimisation to 'simmer'. While 'simmer', as is, can help you answer a question like the following:

> If we have x amount of resources of type A, what will the average waiting time in the process be?

'simmer.optim' is targeted to a reformulation like this:

> What amount x of resources of type A minimises the waiting time, while still maintaining a utilisation level of $\rho_A$?

We would be very grateful if someone with experience on DES optimisation could try it out and give us some feedback. Simply install it from GitHub using 'devtools'

```{r, eval=FALSE}
devtools::install_github("r-simmer/simmer.optim")
```

and start from the [README](https://github.com/r-simmer/simmer.optim), which demonstrates the current functionalities.
