---
title: "simmer 4.0.1"
author: "Iñaki Ucar"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The 4.0.1 release of [simmer](http://r-simmer.org/), the Discrete-Event Simulator for R, is on CRAN since a couple of weeks ago. There are few changes, notably new getters (`get_sources()`, `get_resources()`, `get_trajectory()`) for `simmer` environments and some improvements in resource selection policies (see details in `help(select)`).

A new convenience function, `when_activated`, makes it easier to generate arrivals on demand, triggered from trajectories. Let us consider, for instance, a simple restocking pattern:

```{r}
library(simmer)

restock <- trajectory() %>%
  log_("restock")

serve <- trajectory() %>%
  log_("serve") %>%
  activate("Restock")

env <- simmer() %>%
  add_generator("Customer", serve, at(1, 2, 3)) %>%
  add_generator("Restock", restock, when_activated()) %>%
  run()
```

Finally, this release leverages the [new fast evaluation framework](https://www.enchufa2.es/archives/boost-the-speed-of-r-calls-from-rcpp.html) offered by `Rcpp` (>= 0.12.18) by default, and includes some minor improvements and bug fixes.

### New features:

* New getters (#159):
    * `get_sources()` and `get_resources()` retrieve a character vector of source/resource names defined in a simulation environment.
    * `get_trajectory()` retrieves a trajectory to which a given source is attached.
* New resource selection policies: `shortest-queue-available`, `round-robin-available`, `random-available` (#156). These are the same as the existing non-`available` ones, but they exclude unavailable resources (capacity set to zero). Thus, if all resources are unavailable, an error is raised.

### Minor changes and fixes:

* Rename `-DRCPP_PROTECTED_EVAL` (Rcpp >= 0.12.17.4) as `-DRCPP_USE_UNWIND_PROTECT` (6d27671).
* Keep compilation quieter with `-DBOOST_NO_AUTO_PTR` (70328b6).
* Improve `log_` print (7c2e3b1).
* Add `when_activated()` convenience function to easily generate arrivals on demand from trajectories (#161 closing #160).
* Enhance `schedule` printing (9c66285).
* Fix generator-manager name clashing (#163).
* Deprecate `set_attribute(global=TRUE)`, `get_attribute(global=TRUE)` and `timeout_from_attribute(global=TRUE)` (#164), the `*_global` versions should be used instead.
