---
title: "simmer 3.8.0"
author: "Iñaki Ucar"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The 3.8.0 release of [simmer](http://r-simmer.org/), the Discrete-Event Simulator for R, hit CRAN almost a week ago, and Windows binaries are already available. This version includes two highly requested new features that justify this second consecutive minor release.

### Attachment of precomputed data

Until v3.7.0, the *generator* was the only means to attach data to trajectories, and it was primarily intended for dynamic generation of arrivals:

```{r}
library(simmer)
set.seed(42)

hello_sayer <- trajectory() %>%
  log_("hello!")

simmer() %>%
  add_generator("dummy", hello_sayer, function() rexp(1, 1)) %>%
  run(until=2)
```

Although it may be used to attach precomputed data too, especially using the `at()` adaptor:

```{r}
simmer() %>%
  add_generator("dummy", hello_sayer, at(seq(0, 10, 0.5))) %>%
  run(until=2)
```

Now, let's say that we want to attach some empirical data, and our observations not only include arrival times, but also priorities and some attributes (e.g., measured service times), as in [this question](https://stackoverflow.com/questions/44865924/using-simmer-to-build-a-specific-model) on StackOverflow:

```{r}
myData <- data.frame(
  time = c(1:10,1:5), 
  priority = 1:3, 
  duration = rnorm(15, 50, 5)) %>%
  dplyr::arrange(time)
```

This is indeed possible using generators, but it requires some trickery; more specifically, the clever usage of a *consumer* function as follows:

```{r}
consume <- function(x, prio=FALSE) {
  i <- 0
  function() {
    i <<- i + 1
    if (prio) c(x[[i]], x[[i]], FALSE)
    else x[[i]]
  }
}

activityTraj <- trajectory() %>%
  seize("worker") %>%
  timeout_from_attribute("duration") %>%
  release("worker")

initialization <- trajectory() %>%
  set_prioritization(consume(myData$priority, TRUE)) %>%
  set_attribute("duration", consume(myData$duration)) %>%
  join(activityTraj)

arrivals_gen <- simmer() %>%
  add_resource("worker", 2, preemptive=TRUE) %>%
  add_generator("dummy_", initialization, at(myData$time)) %>%
  run() %>%
  get_mon_arrivals()

# check the resulting duration times
activity_time <- arrivals_gen %>%
  tidyr::separate(name, c("prefix", "n"), convert=TRUE) %>%
  dplyr::arrange(n) %>%
  dplyr::pull(activity_time)

all(activity_time == myData$duration)
```

Since this v3.8.0, the new data source `add_dataframe` greatly simplifies this process:

```{r}
arrivals_df <- simmer() %>%
  add_resource("worker", 2, preemptive=TRUE) %>%
  add_dataframe("dummy_", activityTraj, myData, time="absolute") %>%
  run() %>%
  get_mon_arrivals()

identical(arrivals_gen, arrivals_df)
```

### On-disk monitoring

As some users noted (see [1](https://groups.google.com/d/msg/simmer-devel/A4M7uP0Lmgw/f_0wkctKBQAJ), [2](https://groups.google.com/d/msg/simmer-devel/sO1OlQTpoXw/VL4-_iTBAQAJ)), the default in-memory monitoring capabilities can turn problematic for very long simulations. To address this issue, the `simmer()` constructor gains a new argument, `mon`, to provide different types of monitors. Monitoring is still performed in-memory by default, but as of v3.8.0, it can be offloaded to disk through `monitor_delim()` and `monitor_csv()`, which produce flat delimited files.

```{r}
mon <- monitor_csv()
mon

env <- simmer(mon=mon) %>%
  add_generator("dummy", hello_sayer, function() rexp(1, 1)) %>%
  run(until=2)
env

read.csv(mon$handlers["arrivals"]) # direct access
get_mon_arrivals(env)              # adds the "replication" column
```

See below for a comprehensive list of changes.

### New features:

* New data source `add_dataframe` enables the attachment of precomputed data, in the form of a data frame, to a trajectory. It can be used instead of (or along with) `add_generator`. The most notable advantage over the latter is that `add_dataframe` is able to automatically set attributes and prioritisation values per arrival based on columns of the provided data frame (#140 closing #123).
* New `set_source` activity deprecates `set_distribution()`. It works both for generators and data sources (275a09c, as part of #140).
* New monitoring interface allows for disk offloading. The `simmer()` constructor gains a new argument `mon` to provide different types of monitors. By default, monitoring is performed in-memory, as usual. Additionally, monitoring can be offloaded to disk through `monitor_delim` and `monitor_csv`, which produce flat delimited files. But more importantly, the C++ interface has been refactorised to enable the development of new monitoring backends (#146 closing #119).

### Minor changes and fixes:

* Some documentation improvements (1e14ed7, 194ed05).
* New default `until=Inf` for the `run` method (3e6aae9, as part of #140).
* `branch` and `clone` now accept lists of trajectories, in the same way as `join`, so that there is no need to use `do.call` (#142).
* The argument `continue` (present in `seize` and `branch`) is recycled if only one value is provided but several sub-trajectories are defined (#143).
* Fix process reset: sources are reset in strict order of creation (e7d909b).
* Fix infinite timeouts (#144).
