---
title: "simmer 4.1.0"
author: "Iñaki Ucar"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The 4.1.0 release of [simmer](http://r-simmer.org/), the Discrete-Event Simulator for R, is on CRAN. As per request in the mailing list, now `get_global()` is able to work inside a generator function. Moreover, the new `add_global()` method attaches a global attribute to a simulator.

```{r}
library(simmer)

env <- simmer()

hello_sayer <- trajectory() %>%
  log_("hello world!") %>%
  set_global("interarrival", 1, mod="+")

generator <- function() get_global(env, "interarrival")

env %>%
  add_global("interarrival", 1) %>%
  add_generator("dummy", hello_sayer, generator) %>%
  run(7) %>%
  get_global("interarrival")
```

Compared to plain global variables, these ones are automatically managed and thus reinitialised if the environment is reset.

```{r}
env %>%
  reset() %>%
  get_global("interarrival")

env %>%
  run(7) %>%
  get_global("interarrival")
```

There has been a small refactoring of some parts of the C++ core, which motivates the minor version bump, but this shouldn't be noticeable to the users. Finally, several bug fixes and improvements complete this release. See below for a complete list.

### New features:

* New getter `get_selected()` retrieves names of selected resources via the `select()` activity (#172 addressing #171).
* Source and resource getters have been vectorised to retrieve parameters from multiple entities (as part of #172).
* Simplify C++ `Simulator` interface for adding processes and resources (#162). The responsibility of building the objects has been moved to the caller.
* New `add_global()` method to attach global attributes to a simulation environment (#174 addressing #158).

### Minor changes and fixes:

* Remove 3.8.0 and 4.0.1 deprecations (#170 addressing #165).
* Fix `get_global()` to work outside trajectories (#170 addressing #165).
* Fix `rollback()` with an infinite amount (#173).
* Fix and improve schedules and managers (as part of #174).
* Fix `reset()` to avoid overwriting the simulation environment (#175).
