---
title: "simmer 4.2.1"
author: "Iñaki Ucar"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The 4.2.1 release of [simmer](http://r-simmer.org/), the Discrete-Event Simulator for R, is on CRAN with quite interesting new features and fixes. As discussed in the mailing list, there is a way to handle the specific case in which an arrival is rejected because a queue is full:

```{r}
library(simmer)

reject <- trajectory() %>%
  log_("kicked off...")

patient <- trajectory() %>%
  seize("nurse", continue=FALSE, reject=reject) %>%
  log_("nurse seized") %>%
  timeout(5) %>%
  release("nurse") %>%
  log_("nurse released")

env <- simmer() %>%
  add_resource("nurse", 1, 0) %>%
  add_generator("patient", patient, at(0, 1)) %>%
  run()
```

But as [Tom Lawton pointed out](https://groups.google.com/d/msg/simmer-devel/-KWhDnPuBSQ/U2Ch8U9UCAAJ), until now, there was no way of handling any alternative path for an arrival that was preempted and "kicked off" from a resource. This mechanism has been implemented into the new `handle_unfinished()` activity:

```{r}
patient <- trajectory() %>%
  handle_unfinished(reject) %>%
  seize("nurse") %>%
  log_("nurse seized") %>%
  timeout(5) %>%
  release("nurse") %>%
  log_("nurse released")

env <- simmer() %>%
  add_resource("nurse", 1, 0, preemptive=TRUE, queue_size_strict=TRUE) %>%
  add_generator("LowPrio_patient", patient, at(0), priority=0) %>%
  add_generator("HighPrio_patient", patient, at(1), priority=10) %>%
  run()
```

Note that such a mechanism is more general, because it also covers the first scenario:

```{r}
env <- simmer() %>%
  add_resource("nurse", 1, 0) %>%
  add_generator("patient", patient, at(0, 1)) %>%
  run()
```

Whenever rejection (or preemption) happens and it is catched by the appropriate handler, the new getter `get_seized()` may be useful to know which resource was abandoned.

Finally, the readership may find interesting the new section about the implementation of [state-dependent service rates](https://r-simmer.org/articles/simmer-06-queueing.html#state-dependent-service-rates) in the [Queueing Systems](https://r-simmer.org/articles/simmer-06-queueing.html) vignette. See below for a complete list of changes.

### New features:

- New `handle_unfinished()` activity sets a drop-out trajectory for unfinished
  arrivals, i.e., those dropped from a resource (due to preemption, resource
  shrinkage or a rejected `seize`) or those that `leave` a trajectory (#178
  addressing #177).
- New `release_all()` and `release_selected_all()` activities automatically
  retrieve the amount of resources seized and release it (#180 addressing #25).
- New `get_seized()` and `get_seized_selected()` getters allow an arrival to
  retrieve the amount of resources seized (#180 addressing #179).
- New `stop_if()` activity sets a conditional breakpoint (#181 addressing #100).

### Minor changes and fixes:

- Fix performance issues in data sources (#176).
- Update CITATION.
- Fix monitored activity for preempted arrivals (as part of #178).
- Fix seizes/releases with a null amount (as part of #180).
- Rename internal status codes (as part of #181).
- Provide more context on error or warning (as part of #181).
- Extend the `Queueing Systems` vignette with a section about state-dependent
  service rates.
- Fix performance issues in getters (#183).
